Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2025-01-03 msre.tex
Edits before format review

Chapter \ref{chap:hybrid} laid out the mathematical background and numerical implementation
of the hybrid $S_N$-diffusion method for improving neutronic accuracy in reactor systems containing
highly neutron-absorbing regions. Chapter \ref{chap:hybrid} specifically discussed two transport
correction schemes for the hybrid method: the diffusion and drift correction
schemes. This chapter demonstrates
the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
This chapter employs the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
generate group constants for the hybrid method and to generate reference solutions for method verification.
The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
and to analyze the impact of key input parameters on the hybrid method. The
2-D models test the hybrid method performance through increased spatial dimensionality and
geometric asymmetry relative to the 1-D models. Parametric studies using 2-D models on the $S_N$ angular
discretization reveal its impact on the neutron flux distributions and $k_\text{eff}$.
Lastly, the 3-D models serve as neutronics model
validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}.
This chapter also investigates the computational performance of 3-D simulations on a \gls{HPC}
cluster.

Section \ref{sec:msre-gc} details the OpenMC model setup for group constant generation. Section
\ref{sec:nts-methods} describes and labels the different neutronics methods employed for the
$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details basic test metrics relevant to
neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
\ref{sec:3d-results} discuss the model setups, results, and discussions for the 1-D, 2-D, and 3-D
\gls{MSRE} models, respectively.

@@

Group constants relevant to the neutron diffusion, $S_N$ neutron transport, or both
methods are:
%
\begin{align*}
  \Sigma_{t,g}=& \mbox{ Macroscopic total cross section in group $g$,}\\
  \Sigma_{r,g}=& \mbox{ Macroscopic removal cross section in group $g$,}\\
  \Sigma_s^{g'\rightarrow g}=& \mbox{ Macroscopic group-to-group scattering} \\
  \mbox{cross section matrix,}\\
  \Sigma_{s,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
  group-to-group scattering cross section matrix,}\\
  \Sigma_{sp,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
  group-to-group scattering production} \\
  \mbox{cross section matrix,}\\
  D_g=& \mbox{ $P_1$-based diffusion coefficient in group $g$,}\\
  \nu\Sigma_{f,g}=& \mbox{ Product of the average number of neutrons produced per fission and} \\
  \mbox{the macroscopic fission cross section in group $g$,}\\
  \chi_g=& \mbox{ Neutron fission spectrum in group $g$.}\\
\end{align*}
%
OpenMC generates these group constants using its multigroup cross section generation capability
\cite{boyd_multigroup_2019}. A group constant postprocessing script for Moltres
(\texttt{moltres\_xs.py}) parses OpenMC simulation output files and rewrites the group constant
data into a JSON format file suitable for Moltres simulations. While OpenMC provides most of the
group constants directly, the postprocessing script calculates $\Sigma_{r,g}$ as
%
\begin{align}
  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
  \shortintertext{where}
      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
      group $g$ to $g$.} \nonumber
\end{align}
%
$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
knockout reactions into the scattering cross section. Neutronics codes commonly identify neutron
knockout reactions as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which
to incorporate neutron knockout effects in the neutron diffusion method.
All simulations in this thesis used 2nd-order-accurate scattering cross sections approximated using
Legendre polynomials.

@@

Material compositions and densities of all components in the \gls{MSRE} models are from reference
\gls{MSRE} specifications \cite{robertson_msre_1965, fratoni_molten_2020}.
All OpenMC simulations in this work used the ENDF/B-VII.1 nuclear data library
\cite{chadwick_endf/b-vii.1_2011}. All models have a uniform
temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
1-D, 2-D, and 3-D OpenMC models may be found in the respective subsections below.

@@

I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
values to assess the accuracy of the deterministic methods ($S_N$, neutron diffusion, hybrid 
methods) against OpenMC in continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
\ref{table:var} details how OpenMC and the deterministic methods represent the position, direction
of travel, neutron energy, and angle-dependence in $\Sigma_s$. The hybrid $S_N$-diffusion method
features representations from either $S_N$ or neutron diffusion depending on the subregion under
consideration. Comparing OpenMC-MG results with OpenMC-CE results allows us
to quantify errors arising from neutron energy group discretization and the 2nd-order scattering
cross section approximations.

@@

\begin{table}[htb!]
  \centering
  \footnotesize
  \caption{Implementation differences of neutronics methods on Moltres and Python.}
  \begin{tabular}{p{0.14\linewidth} p{0.25\linewidth} p{0.15\linewidth} p{0.15\linewidth} p{0.17\linewidth}}
    \toprule
    Implementations & Spatial discretization & $S_N$ formulation & $S_N$ acceleration &
    Hybrid transport\newline correction \\
    \midrule
    Moltres & Finite element method & SAAF & Nonlinear diffusion acceleration & Drift correction \\
    Python & Finite difference (Diffusion)\newline Diamond difference ($S_N$) & First-order & None & Diffusion correction \\
    \bottomrule
  \end{tabular}
  \label{table:nts-methods}
\end{table}

\section{Basic Test Metrics} \label{sec:test-metrics}

This section defines several metrics for comparing simulation results among the various
neutronics methods.

For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
multiplication factor ($k_\text{eff}$), the rate of neutron population increase between successive
neutron generations. Reactivity is defined as
%
\begin{gather}
  \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
\end{gather}
%
Differences in reactivity, for either comparisons between neutronics methods or control rod worth
calculations, can be calculated as
%
\begin{gather}
  \Delta\rho_{1,2} = \rho_1 - \rho_2 =
  \frac{k_{\text{eff},1}-k_{\text{eff},2}}{k_{\text{eff},1}k_{\text{eff},2}}.
\end{gather}

@@

This section presents the models, results, and discussion for the 1-D test cases.

\subsection{1-D Neutronics Model Setup \& Geometries}

@@

\subsection{1-D Neutronics Model Mesh Convergence Tests}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=.95\columnwidth]{diffusion-mesh-convergence-flux}
  \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
  mesh refinement levels 1 to 3 relative to the reference group flux distributions at mesh
  refinement level 4. Flux differences fall by an order of magnitude with each successive mesh
  refinement.}
  \label{fig:diff-mesh-flux}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=.95\columnwidth]{sn-mesh-convergence-flux}
  \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
  for mesh refinement levels 1 to 3 relative to the reference group flux distributions at mesh
  refinement level 4. Flux differences fall by an order of magnitude with each successive mesh
  refinement.}
  \label{fig:sn-mesh-flux}
\end{figure}

@@

\subsection{1-D Neutronics Numerical Results \& Discussion}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{rho}
  \caption{Difference in reactivity $\rho$ of all neutronics methods investigated relative
  to OpenMC-CE.}
  \label{fig:1d-rho}
\end{figure}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{worth}
  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
  investigated relative to OpenMC-CE.}
  \label{fig:1d-worth}
\end{figure}

@@

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3a-flux}
  \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
  \label{fig:3a-flux}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-flux}
  \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
  \label{fig:3b-flux}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3a-flux-diff}
  \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
  \label{fig:3a-flux-diff}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-flux-diff}
  \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
  \label{fig:3b-flux-diff}
\end{figure}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-drift}
  \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
  Moltres-hybrid and Moltres-$S_8$ solvers.}
  \label{fig:3b-drift}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-svdc}
  \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
    Python-hybrid and Python-$S_8$ solvers.}
  \label{fig:3b-svdc}
\end{figure}

@@

The hybrid $S_N$-diffusion method relies on $S_N$ calculations in the correction subregion to
generate flux corrections. Minimizing the size of the correction subregion and the $S_N$ subproblem
is essential for the hybrid method to be computationally competitive for time-dependent full-core
simulations. I investigated the effect of the correction subregion size on the $k_\text{eff}$ and
control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
at 5 cm-intervals.

@@

\begin{table}[htb]
  \centering
  \caption{Number of outer iterations in hybrid method calculations of Case 3b for a given set of
  convergence tolerance values imposed on the $S_8$ subsolver.}
  \begin{tabular}{l S S S S S S}
    \toprule
    $S_8$ subsolver tolerance, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
    \midrule
    Number of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
    \bottomrule
  \end{tabular}
  \label{table:sn-tol}
\end{table}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=0.6\columnwidth]{sn-tol}
  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
  convergence.}
  \label{fig:sn-tol}
\end{figure}

@@

This section presents the models, results, and discussion of the 2-D cases.

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{quarter-core-geom}
  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
  \gls{MSRE} geometry.}
  \label{fig:1/4-geom}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{full-core-geom}
  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
  \gls{MSRE} geometry.}
  \label{fig:full-geom}
\end{figure}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
    model.}
  \label{fig:full-geom-closeup}
\end{figure}

@@

@@

Table \ref{table:quarter-core} details the $k_\text{eff}$ and control rod worth estimates and error
values relative to OpenMC-CE. For the ``No Rod'' case, OpenMC-CE estimated the $k_\text{eff}$ to be
1.11209(43). The OpenMC-MG, hybrid, and neutron diffusion methods performed
similarly to one another with reactivity errors ranging from 618 pcm to 773 pcm relative to
OpenMC-CE. Based on $k_\text{eff}$ alone, the hybrid method performed marginally worse than the
neutron diffusion method. However, for the ``Rod'' case, then neutron diffusion method saw a
change in sign of the error to -816 pcm. The errors for OpenMC-MG and the hybrid method
remained positive and reduced to 446 pcm and 760 pcm, respectively. Consequently, when taking the
difference in reactivities to estimate the rod worth, the neutron diffusion method performed
significantly worse than OpenMC-MG and the hybrid method. OpenMC-CE estimated the rod worth to be
8370(53) pcm. The neutron diffusion method overestimated the rod worth by 1484 pcm. The hybrid
method improved significantly over the neutron diffusion method with a rod worth estimate that is
only 13 pcm higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm. This is likely due
to statistical uncertainty in OpenMC-MG and favorable error cancellation between different sources
of error in the hybrid method (e.g., neutron leakage, heterogeneous core geometry).
.
The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.

@@

This section presents the models, results, and discussion of the 3-D \gls{MSRE} simulations.

@@

\begin{table}[htb]
  \centering
  \caption{List of key reactor dimensions.}
  \begin{tabular}{l S}
    \toprule
    Reactor Specifications & {Value} \\
    \midrule
    Total height [cm] & 215.5967 \\
    Lower plenum height [cm] & 18.7530 \\
    Upper plenum height [cm] & 25.3937 \\
    Total radius [cm] & 76.200 \\
    Graphite core radius [cm] & 70.168 \\
    Core can inner radius [cm] & 70.485 \\
    Core can outer radius [cm] & 71.120 \\
    Vessel thickness [cm] & 2.540 \\
    Fuel channel length [cm] & 3.048 \\
    Fuel channel width [cm] & 1.016 \\
    Graphite lattice pitch [cm] & 5.080 \\
    Control rod inner radius [cm] & 1.0668 \\
    Control rod outer radius [cm] & 1.3716 \\
    Control rod thimble inner radius [cm] & 2.3749 \\
    Control rod thimble outer radius [cm] & 2.540 \\
    Homogenized sample basket radius [cm] & 2.8575 \\
    Radius of fuel annulus around thimble [cm] & 3.2016 \\
    \bottomrule
  \end{tabular}
  \label{table:reactor-dimensions}
\end{table}

@@

Figures * show a 3-D section view of the \gls{MSRE} model geometry and the group 1 \& 8 neutron
flux distributions. The salt-graphite lattice structure is visible from the group flux
distributions. The group 1 flux distribution largely retains a spherical shape. The group 8 neutron
flux distribution is particularly suppressed near the control rods due to strong thermal neutron
capture cross section of gadolinium.

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{ic-g1}
  \caption{}
  \label{fig:ic-g1}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{ic-g8}
  \caption{}
  \label{fig:ic-g8}
\end{figure}

@@

Analysis on the drift and diffusion correction schemes showed that the diffusion correction
parameter led to numerous discontinuities tending to infinity values near points in the problem
domain where flux peaks and troughs occur. These discontinuities and negative diffusion coefficient
values are unphysical and would lead to numerical stability issues. Therefore, the hybrid method
implementation with drift correction scheme was chosen for the remaining 1-D analyses and the
2-D and 3-D analyses.

>>>>

Chapter \ref{chap:hybrid} laid out the mathematical background and numerical implementation
of the hybrid $S_N$-diffusion method for improving neutronic accuracy in reactor systems containing
highly neutron-absorbing regions. Chapter \ref{chap:hybrid} specifically discussed two transport
correction schemes for the hybrid method: the diffusion and drift correction
schemes. This chapter demonstrates the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue
simulations of the \gls{MSRE}. The work in this chapter employs the OpenMC Monte Carlo neutral
particle transport software \cite{romano_openmc:_2015} to generate group constants for the hybrid
method and to generate reference solutions for method verification.
The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
and to analyze the impact of key input parameters on the hybrid method. The
2-D models test the hybrid method performance through increased spatial dimensionality and
geometric asymmetry relative to the 1-D models. Parametric studies using 2-D models on the $S_N$ angular
discretization reveal its impact on the neutron flux distributions and $k_\text{eff}$.
Lastly, the 3-D models serve as neutronics model
validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}.
This chapter also investigates the computational performance of 3-D simulations on a \gls{HPC}
cluster.

Section \ref{sec:msre-gc} details the OpenMC model setup for group constant generation. Section
\ref{sec:nts-methods} describes and labels the different neutronics methods employed for the
$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details test metrics relevant to
neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
\ref{sec:3d-results} discuss the model setups, results, and discussions for the 1-D, 2-D, and 3-D
\gls{MSRE} models, respectively.

@@

Deterministic neutron diffusion and $S_N$ neutron transport methods require group constant data in
the form of group-averaged macroscopic neutron cross sections, neutron speeds, and delayed neutron
precursor data.
Group constants relevant to the neutron diffusion, $S_N$ neutron transport, or both
methods in Moltres are:
%
\begin{align*}
  \Sigma_{t,g}=& \mbox{ Macroscopic total cross section in group $g$,} &\\
  \Sigma_{r,g}=& \mbox{ Macroscopic removal cross section in group $g$,} &\\
  \Sigma_s^{g'\rightarrow g}=& \mbox{ Macroscopic group-to-group scattering cross section matrix,} &\\
  \Sigma_{s,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
        group-to-group scattering} & \\
                           & \mbox{cross section matrix,} &\\
  \Sigma_{sp,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
  group-to-group scattering production} & \\
                                        & \mbox{cross section matrix,} &\\
  D_g=& \mbox{ $P_1$-based diffusion coefficient in group $g$,} &\\
  \nu\Sigma_{f,g}=& \mbox{ Product of the average number of neutrons produced per fission and} &\\
                  &\mbox{the macroscopic fission cross section in group $g$,}&\\
  \chi_g=& \mbox{ Neutron fission spectrum in group $g$.}&\\
\end{align*}

OpenMC generates these group constants using its multigroup cross section generation capability
\cite{boyd_multigroup_2019}. A group constant postprocessing script for Moltres
(\texttt{moltres\_xs.py}) parses OpenMC simulation output files and rewrites the group constant
data into a JSON format file suitable for Moltres simulations. While OpenMC provides most of the
group constants directly, the postprocessing script calculates $\Sigma_{r,g}$ as
%
\begin{align}
  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
  \shortintertext{where}
      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
      group $g$ to $g$.} \nonumber
\end{align}
%
$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
knockout reactions into the scattering cross section. Neutronics codes commonly identify neutron
knockout reactions as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which
to incorporate neutron knockout effects in the neutron diffusion method.
All simulations in this thesis used 2nd-order-accurate scattering cross sections approximated using
Legendre polynomials.

@@

Material compositions and densities of all components in the \gls{MSRE} models are from reference
\gls{MSRE} specifications \cite{robertson_msre_1965, fratoni_molten_2020}.
All OpenMC simulations in this work used the ENDF/B-VII.1 nuclear data library
\cite{chadwick_endf/b-vii.1_2011}. All models have a uniform
temperature of 900 K unless otherwise stated. Sections \ref{sec:1d-model-setup},
\ref{sec:2d-model-setup}, \ref{sec:3d-model-setup} provide detailed information pertaining
specifically to the 1-D, 2-D, and 3-D model geometries, respectively.

@@

Besides generating group constants for the deterministic methods, OpenMC also provides reference
netruon flux and $k_\text{eff}$ values for all studied cases. OpenMC can run in continuous-energy
(OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
\ref{table:var} details how OpenMC and the deterministic methods represent the position, direction
of travel, neutron energy, and angle-dependence in $\Sigma_s$. The hybrid $S_N$-diffusion method
features representations from either $S_N$ or neutron diffusion depending on the subregion under
consideration. Comparing OpenMC-MG results with OpenMC-CE results allows us
to quantify errors arising from neutron energy group discretization and the 2nd-order scattering
cross section approximations.

@@

\begin{table}[htb!]
  \centering
  \footnotesize
  \caption{Implementation differences of neutronics methods on Moltres and Python.}
  \begin{tabular}{l l l}
    \toprule
    Implementation detail & Moltres & Python \\
    \midrule
    Spatial discretization & Finite element method & Finite difference \& Diamond difference \\
    $S_N$ formulation & SAAF & First-order \\
    $S_N$ acceleration & Nonlinear diffusion acceleration & None \\
    Hybrid transport correction & Drift correction & Diffusion correction \\
    \bottomrule
  \end{tabular}
  \label{table:nts-methods}
\end{table}

\section{Test Metrics} \label{sec:test-metrics}

This section defines several metrics for comparing simulation results among the various
neutronics methods.
For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
multiplication factor ($k_\text{eff}$), the rate of neutron population increase between successive
neutron generations. Reactivity is defined as
%
\begin{gather}
  \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
\end{gather}
%
Differences in reactivity, for either comparisons between neutronics methods or control rod worth
calculations, can be calculated as
%
\begin{gather}
  \Delta\rho_{1,2} = \rho_1 - \rho_2 =
  \frac{k_{\text{eff},1}-k_{\text{eff},2}}{k_{\text{eff},1}k_{\text{eff},2}}.
\end{gather}

@@

This section presents the models, results, and discussion for the 1-D test cases. Section
\ref{sec:1d-model-setup} describes the model setup and geometries of the 1-D test cases.
Section \ref{sec:1d-mesh-conv} discusses the 1-D mesh convergence tests. Lastly, Section
\ref{sec:1d-results-sub} shows the results and discussion of the 1-D test cases.

\subsection{1-D Neutronics Model Setup \& Geometries} \label{sec:1d-model-setup}

@@

\subsection{1-D Neutronics Model Mesh Convergence Tests} \label{sec:1d-mesh-conv}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=.96\columnwidth]{diffusion-mesh-convergence-flux}
  \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
  mesh refinement levels 1 to 3 relative to the reference group flux distributions at mesh
  refinement level 4. Flux differences fall by an order of magnitude with each successive mesh
  refinement.}
  \label{fig:diff-mesh-flux}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=.96\columnwidth]{sn-mesh-convergence-flux}
  \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
  for mesh refinement levels 1 to 3 relative to the reference group flux distributions at mesh
  refinement level 4. Flux differences fall by an order of magnitude with each successive mesh
  refinement.}
  \label{fig:sn-mesh-flux}
\end{figure}

@@

\subsection{1-D Neutronics Numerical Results \& Discussion} \label{sec:1d-results-sub}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{rho}
  \caption{Difference in reactivity $\rho$ of all neutronics methods investigated relative
  to OpenMC-CE. The standard deviation refers to the uncertainty associated with the Monte Carlo
  method in OpenMC-CE.}
  \label{fig:1d-rho}
\end{figure}

@@

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\columnwidth]{worth}
  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
  investigated relative to OpenMC-CE. All percentage difference values are consistently positive.}
  \label{fig:1d-worth}
\end{figure}

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{case-3a-flux}
  \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
  \label{fig:3a-flux}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-flux}
  \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
  \label{fig:3b-flux}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{case-3a-flux-diff}
  \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
  \label{fig:3a-flux-diff}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-flux-diff}
  \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
  \label{fig:3b-flux-diff}
\end{figure}

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=.99\columnwidth]{case-3b-drift}
  \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
  Moltres-hybrid and Moltres-$S_8$ solvers.}
  \label{fig:3b-drift}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=.99\columnwidth]{case-3b-svdc}
  \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
    Python-hybrid and Python-$S_8$ solvers.}
  \label{fig:3b-svdc}
\end{figure}

@@

The hybrid $S_N$-diffusion method relies on $S_N$ calculations in the correction subregion to
generate flux corrections. Minimizing the size of the correction subregion and the $S_N$ subproblem
is essential for the hybrid method to be computationally competitive for time-dependent full-core
simulations. I investigated the effect of the correction subregion size on the $k_\text{eff}$ and
control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
at 5 cm-intervals. All subregion sizes place the subregion boundary more than 1 neutron mean free
path away from the control rod. The shortest measured neutron mean free paths are 4.81 cm and 5.85
cm for group 1 neutrons in the fuel and graphite regions, respectively.

@@

\begin{table}[h]
  \centering
  \caption{Number of outer iterations in hybrid method calculations of Case 3b for a given set of
  convergence tolerance values imposed on the $S_8$ subsolver.}
  \begin{tabular}{l S S S S S S}
    \toprule
    $S_8$ subsolver tolerance, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
    \midrule
    Number of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
    \bottomrule
  \end{tabular}
  \label{table:sn-tol}
\end{table}
%
\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\columnwidth]{sn-tol}
  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
  convergence.}
  \label{fig:sn-tol}
\end{figure}

@@

@@

This section presents the models, results, and discussion of the 2-D cases. Section
\ref{sec:2d-model-setup} describes the model setup of the 2-D cases. Section
\ref{sec:2d-nts-results} covers the results and discussion of the 2-D cases.

@@

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=\columnwidth]{quarter-core-geom}
  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
  \gls{MSRE} geometry.}
  \label{fig:1/4-geom}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{full-core-geom}
  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
  \gls{MSRE} geometry.}
  \label{fig:full-geom}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
    model.}
  \label{fig:full-geom-closeup}
\end{figure}

@@

Table \ref{table:quarter-core} details the $k_\text{eff}$ and control rod worth estimates and error
values relative to OpenMC-CE. For the ``No Rod'' case, OpenMC-CE estimated the $k_\text{eff}$ to be
1.11209(43). The OpenMC-MG, hybrid, and neutron diffusion methods performed
similarly to one another with reactivity errors ranging from 618 pcm to 773 pcm relative to
OpenMC-CE. Based on $k_\text{eff}$ alone, the hybrid method performed marginally worse than the
neutron diffusion method. However, for the ``Rod'' case, then neutron diffusion method saw a
change in sign of the error to -816 pcm. The errors for OpenMC-MG and the hybrid method
remained positive and reduced to 446 pcm and 760 pcm, respectively. Consequently, when taking the
difference in reactivities to estimate the rod worth, the neutron diffusion method performed
significantly worse than OpenMC-MG and the hybrid method. OpenMC-CE estimated the rod worth to be
8370(53) pcm. The neutron diffusion method overestimated the rod worth by 1484 pcm. The hybrid
method improved significantly over the neutron diffusion method with a rod worth estimate that is
only 13 pcm higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm. This is likely due
to statistical uncertainty in OpenMC-MG and favorable error cancellation between different sources
of error in the hybrid method (e.g., neutron leakage, heterogeneous core geometry).
The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.

@@

This section presents the models, results, and discussion of the 3-D \gls{MSRE} simulations.
Section \ref{sec:3d-model-setup} describes the model setup for the 3-D \gls{MSRE} simulations.
Section \ref{sec:3d-nts} covers the results and discussion of the 3-D \gls{MSRE} simulations.

@@

\begin{table}[htb]
  \centering
  \caption{List of key dimensions of the 3-D \gls{MSRE} model used in this work.}
  \begin{tabular}{l S}
    \toprule
    Reactor Specifications & {Value} \\
    \midrule
    Total height [cm] & 215.5967 \\
    Lower plenum height [cm] & 18.7530 \\
    Upper plenum height [cm] & 25.3937 \\
    Total radius [cm] & 76.200 \\
    Graphite core radius [cm] & 70.168 \\
    Core can inner radius [cm] & 70.485 \\
    Core can outer radius [cm] & 71.120 \\
    Vessel thickness [cm] & 2.540 \\
    Fuel channel length [cm] & 3.048 \\
    Fuel channel width [cm] & 1.016 \\
    Graphite lattice pitch [cm] & 5.080 \\
    Control rod inner radius [cm] & 1.0668 \\
    Control rod outer radius [cm] & 1.3716 \\
    Control rod thimble inner radius [cm] & 2.3749 \\
    Control rod thimble outer radius [cm] & 2.540 \\
    Homogenized sample basket radius [cm] & 2.8575 \\
    Radius of fuel annulus around thimble [cm] & 3.2016 \\
    \bottomrule
  \end{tabular}
  \label{table:reactor-dimensions}
\end{table}

@@

Figures \ref{fig:3d-view}, \ref{fig:ic-g1}, and \ref{fig:ic-g8} show a 3-D section view of the
\gls{MSRE} model geometry and the group 1 \& 8 neutron
flux distributions. The salt-graphite lattice structure is visible from the group flux
distributions. The group 1 flux distribution largely retains a spherical shape. The group 8 neutron
flux distribution is particularly suppressed near the control rods due to strong thermal neutron
capture cross section of gadolinium.

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{ic-g1}
  \caption{Group 1 neutron flux distribution with Rod 1 inserted by 4.4 inches at initial
  criticality.}
  \label{fig:ic-g1}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{ic-g8}
  \caption{Group 8 neutron flux distribution with Rod 1 inserted by 4.4 inches at initial
  criticality.}
  \label{fig:ic-g8}
\end{figure}

@@

Analysis on the drift and diffusion correction schemes showed that the diffusion correction
parameter led to numerous discontinuities tending to infinity values near points in the problem
domain where flux peaks and troughs occur. These discontinuities and negative diffusion coefficient
values are unphysical and lead to numerical stability issues. Therefore, the hybrid method
implementation with drift correction scheme was chosen for the remaining 1-D analyses and the
2-D and 3-D analyses.
